---
title: "MTH 404 R Project -- some code"
date: "2023-04-12"
output:
  pdf_document: default
  world_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Loading Packages

```{r}
library(tidyverse)
library(corrplot)
library(ggplot2)
library(lubridate)
library(gridExtra)
library(caTools)
library(GGally)
```

## Reading Data

```{r}
mydata=read.csv("/Users/ashwinram/Documents/MATLAB/train.csv")
test_data = read.csv("/Users/ashwinram/Documents/MATLAB/test.csv")
```

Data structure and summary statistics
```{r}
str(mydata)
summary(mydata)
```

Removing Qualitative Variables

```{r echo = T, results = 'hide'}
len = length(mydata[1,])
counter = 0
while(counter != len){
  for (x in 1:len) {
    eval.parent(substitute(counter <- counter + 1))
    if(counter == length(mydata[1,]) + 1){
      counter = len
      break
    }
    if(sapply(mydata[1,x],class) == "character"){
      counter = 0
      mydata = mydata[,-x]
      print(x)
      print(mydata[,-x])
      break
    }
  }
}
len = length(test_data[1,])
counter = 0
while(counter != len){
  for (x in 1:len) {
    eval.parent(substitute(counter <- counter + 1))
    if(counter == length(test_data[1,]) + 1){
      counter = len
      break
    }
    if(sapply(test_data[1,x],class) == "character"){
      counter = 0
      test_data = test_data[,-x]
      print(x)
      print(test_data[,-x])
      break
    }
  }
}
len = length(test_data[1,])
for(x in test_data[,1]){
  for(y in 1:len){
    if(is.na(test_data[x,y])){
      test_data[x,y] = 0
    }  
  }
}
```

Replacing NA Values with 0

```{r}
len = length(mydata[1,])
for(x in mydata[,1]){
  for(y in 1:len){
    if(is.na(mydata[x,y])){
      mydata[x,y] = 0
    }  
  }
}
len = length(test_data[1,])

```

## Data visualization 

Check the distribution of the target variable SalePrice

```{r}
boxplot(mydata$SalePrice)
hist (mydata$SalePrice)
```

### Determining the association between variables.

We take out the correlation plot (corrplot) to understand the association of the dependent variable (price) with the independent variables.
* numerical variable only

```{r}
cor_data=data.frame(mydata[,2:38])
correlation=cor(cor_data)
par(mfrow=c(1, 1))
corrplot(correlation,method="color")
```

Graphing scatter plots of the variables that highly correlated to the SalePrice

```{r echo = T, results = 'hide'}
P1 = ggplot(data = mydata, aes(x = YearBuilt, y = SalePrice)) +
  geom_point() + geom_smooth(method = "lm", se = FALSE)
P2 = ggplot(data = mydata, aes(x = YearRemodAdd, y = SalePrice)) +
  geom_point() + geom_smooth(method = "lm", se = FALSE)
P4 = ggplot(data = mydata, aes(x = BsmtUnfSF, y = SalePrice)) +
  geom_point() + geom_smooth(method = "lm", se = FALSE)
P5 = ggplot(data = mydata, aes(x = TotalBsmtSF, y = SalePrice)) +
  geom_point() + geom_smooth(method = "lm", se = FALSE)
P6 = ggplot(data = mydata, aes(x = GrLivArea, y = SalePrice)) +
  geom_point() + geom_smooth(method = "lm", se = FALSE)
P11 = ggplot(data = mydata, aes(x = GarageArea, y = SalePrice)) +
  geom_point() + geom_smooth(method = "lm", se = FALSE)
P12 = ggplot(data = mydata, aes(x = LotArea, y = SalePrice)) +
  geom_point() + geom_smooth(method = "lm", se = FALSE)
P13 = ggplot(data = mydata, aes(x = X1stFlrSF , y = SalePrice)) +
  geom_point() + geom_smooth(method = "lm", se = FALSE)
P14 = ggplot(data = mydata, aes(x = X2ndFlrSF , y = SalePrice)) +
  geom_point() + geom_smooth(method = "lm", se = FALSE)
P15 = ggplot(data = mydata, aes(x =BsmtFinSF1 , y = SalePrice)) +
  geom_point() + geom_smooth(method = "lm", se = FALSE)
P18 = ggplot(data = mydata, aes(x =WoodDeckSF , y = SalePrice)) +
  geom_point() + geom_smooth(method = "lm", se = FALSE)
P19 = ggplot(data = mydata, aes(x =OpenPorchSF , y = SalePrice)) +
  geom_point() + geom_smooth(method = "lm", se = FALSE)

```

Graphing Boxplots of the variables that highly correlated to the SalePrice

```{r}
par(mfrow=c(1, 5))
boxplot(SalePrice~OverallQual,data=mydata,main="Different boxplots", xlab="OverallQual",ylab="SalePrice",col="red",border="orange")
boxplot(SalePrice~FullBath,data=mydata,main="Different boxplots", xlab="view",ylab="SalePrice",col="red",border="orange")
boxplot(SalePrice~TotRmsAbvGrd,data=mydata,main="Different boxplots", xlab="TotRmsAbvGrd",ylab="SalePrice",col="red",border="orange")
boxplot(SalePrice~Fireplaces,data=mydata,main="Different boxplots", xlab="Fireplaces",ylab="SalePrice",col="red",border="orange")
boxplot(SalePrice~GarageCars,data=mydata,main="Different boxplots", xlab="GarageCars",ylab="SalePrice",col="red",border="orange")
```

#### Modifying Train Data

 Adding a variable that represents the age of the house to train data
 
```{r}
mydata$Sale = mydata$YrSold + mydata$MoSold/12
mydata$Age = -(mydata$Sale - mydata$YearBuilt)
test_data$Sale = test_data$YrSold + test_data$MoSold/12
test_data$Age = -(test_data$Sale - test_data$YearBuilt)
```

Removing Outliers

```{r}
outliers=boxplot(mydata$SalePrice,plot=FALSE)$out
outliers_data=mydata[which(mydata$price %in% outliers),]
mydata1= mydata[-which(mydata$SalePrice %in% outliers),]
```

##### Modeling Train Data

Creating a model based on the variables the correlated the most to Train Data
```{r}
model=lm(data=mydata1,SalePrice~YearBuilt+YearRemodAdd+BsmtUnfSF+TotalBsmtSF+GarageArea+LotArea+X1stFlrSF+X2ndFlrSF+WoodDeckSF+OpenPorchSF+OverallQual+TotRmsAbvGrd+Fireplaces+GarageCars)
summary(model)
```

Detecting Influential Points and removing them

```{r echo = T, results = 'hide'}
cooksd <- cooks.distance(model)
cooksd
mean(cooksd)
influential <- as.numeric(names(cooksd)[(cooksd > 4*mean(cooksd, na.rm=T))])
influential_data=mydata1[influential, ]
influencial_outliers=inner_join(outliers_data,influential_data)
mydata12=rbind(mydata1,influencial_outliers)
```

Modeling the new Train Data

```{r}
model=lm(data=mydata12,SalePrice~YearBuilt+YearRemodAdd+BsmtUnfSF+TotalBsmtSF+GarageArea+LotArea+X1stFlrSF+X2ndFlrSF+WoodDeckSF+OpenPorchSF+OverallQual+TotRmsAbvGrd+Fireplaces+GarageCars)
summary(model)
```

Accuracy of the Model

```{r echo = T, results = 'hide'}
pred=model$fitted.values
pred
tally_table=data.frame(actual=mydata12$SalePrice, predicted=pred)
tally_table
mape=mean(abs(tally_table$actual-tally_table$predicted)/tally_table$actual)
accuracy=1-mape
accuracy

```
Predicting Test Sale Price

```{r}
test_data_1=test_data[,c(5,7,8,12,13,28,14,15,4,29,30,24,25,27)]
pred_test=predict(newdata=test_data_1,model)
Id <- c(1460:2919)
SalePrice<- c()
for(x in 1:size){
  SalePrice <- append(SalePrice,pred_test[[x]])
}
SalePrice
df <- data.frame(Id, SalePrice) 
df
```

Creating File containing estimated prices

```{r}
write.csv(df, "Ptaeople.csv", row.names=FALSE)
```


